---
title: "Blog Examples"
author: "Theo Roe"
date: "4 October 2017"
output: html_document
---
<style>
    body .main-container {
        max-width: 600px;
    }
    .column-left{
  float: left;
  width: 50%;
  text-align: left;
}
.column-center{
  display: inline-block;
  width: 33%;
  text-align: center;
}
.column-right{
  float: right;
  width: 50%;
  text-align: left;
}

  .col2 {
    columns: 2 200px;         /* number of columns and width in pixels*/
    -webkit-columns: 2 200px; /* chrome, safari */
    -moz-columns: 2 200px;    /* firefox */
  }
  .col3 {
    columns: 3 100px;
    -webkit-columns: 3 100px;
    -moz-columns: 3 100px;
  }

</style>
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(tidyverse)
```



1) __*Creation.*__ Just as `data.frame` creates data frames,`tibble`, `data_frame` and `tribble` all create tibbles. 

<div class="col2">
Standard data frame
```{r}
data.frame(a = 1:5, 
           b = letters[1:5])  
```
tibble using `data_frame`
```{r}
data_frame(a = 1:5, 
           b = letters[1:5])
```
tibble using `tibble`
```{r}
tibble(a = 1:5, 
           b = letters[1:5])
```
tibble using `tribble`
```{r}
tribble( ~a, ~b, 
       #---|----
          1, "a", 
          2, "b")
```
</div>

2) __*Coercion.*__ Just as `as.data.frame` coerces objects into data frames, `as_data_frame` and `as_tibble` coerce objects into tibbles. 
```{r}
df = data.frame(a = 1:5, b = letters[1:5])
```
<div class="col2">
```{r}
as_data_frame(df)
as_tibble(df)
```
</div>

3) __*Importing.*__ Just as `read.csv`

[click here](tibbles.html)