---
title: "How to Form a Tibble"
author: "Theo Roe"
date: "2 October 2017"
output: html_document
---

[<- Back to the main page](tibbles.html)

<style>
    body .main-container {
        max-width: 600px;
    }
    .column-left{
  float: left;
  width: 50%;
  text-align: left;
}
.column-center{
  display: inline-block;
  width: 33%;
  text-align: center;
}
.column-right{
  float: right;
  width: 50%;
  text-align: left;
}

  .col2 {
    columns: 2 200px;         /* number of columns and width in pixels*/
    -webkit-columns: 2 200px; /* chrome, safari */
    -moz-columns: 2 200px;    /* firefox */
  }
  .col3 {
    columns: 3 100px;
    -webkit-columns: 3 100px;
    -moz-columns: 3 100px;
  }

</style>
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(tidyverse)
```

Here are those examples you were looking for. 

1) __*Creation.*__ Just as `data.frame` creates data frames,`tibble`, `data_frame` and `tribble` all create tibbles. 

<div class="col2">
Standard data frame.
```{r}
data.frame(a = 1:5, 
           b = letters[1:5])  
```
A tibble using `data_frame`.
```{r}
data_frame(a = 1:5, 
           b = letters[1:5])
```
A tibble using `tibble`.
```{r}
tibble(a = 1:5, 
           b = letters[1:5])
```
A tibble using `tribble`.
```{r}
tribble( ~a, ~b, 
       #---|----
          1, "a", 
          2, "b")
```
</div>
Notice the odd one out? Tribbles are different. They're a way of laying out small amounts of data in an easy to read form. I'm not too keen on these, as even writing out that simple 2 x 2 tribble got tedious. 

2) __*Coercion.*__ Just as `as.data.frame` coerces objects into data frames, `as_data_frame` and `as_tibble` coerces objects into tibbles. 
```{r}
df = data.frame(a = 1:5, b = letters[1:5])
```
<div class="col2">
```{r}
as_data_frame(df)
as_tibble(df)
```
</div>

You can coerce more than just data frames, too. These take lists, matrices, vectors and single instances of class. 

3) __*Importing.*__ There's a few options to read in data files within the `tidyverse`, so we'll just compare `read_csv` and its representative `data.frame` pal, `read.csv`. Let's take a look at them. 
I'm going to use the same csv file as in the main page.
```{r, include=FALSE}
tib = read_csv("x.csv")
options(tibble.print_max = 10, tibble.print_min = 5)
```
<div class = "col2">
```{r, eval = FALSE}
tib = read_csv("x.csv")
```
```{r}
tib
```

***

```{r}
df = read.csv("x.csv")
```
```{r}
df
```
</div>
Not only does `read_csv` return the pretty tibble, it is also much faster. For proof, check out [this article](http://yetanothermathprogrammingconsultant.blogspot.co.uk/2016/12/reading-csv-files-in-r-readcsv-vs.html) by Erwin Kalvelagen. 