---
title: "The Trouble with Tibbles"
author: "Theo Roe"
date: "2 October 2017"
output: html_document
---
<style>
    body .main-container {
        max-width: 600px;
    }
    .column-left{
  float: left;
  width: 50%;
  text-align: left;
}
.column-center{
  display: inline-block;
  width: 33%;
  text-align: center;
}
.column-right{
  float: right;
  width: 50%;
  text-align: left;
}

  .col2 {
    columns: 2 200px;         /* number of columns and width in pixels*/
    -webkit-columns: 2 200px; /* chrome, safari */
    -moz-columns: 2 200px;    /* firefox */
  }
  .col3 {
    columns: 3 100px;
    -webkit-columns: 3 100px;
    -moz-columns: 3 100px;
  }

</style>
```{r setup,include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Let's get something straight, there isn't really any trouble with tibbles. I'm hoping you've noticed this is a play on 1967 Star Trek episode, "The Trouble with Tribbles". I've recently got myself a job as a Data Scientist, here, at [Jumping Rivers](https://www.jumpingrivers.com/). Having never come across tibbles until this point, I now find myself using them in nearly every R script I compose. Be that your timeless standard R script, your friendly Shiny app or an analytical Markdown document. 

### _What are tibbles?_

Presumably this is why you came here, right?

Tibbles ARE data frames. *Shocked emoji*. Yep, they are frames of data. There's a quote I found somewhere on the internet that decribes tibbles quite well;  

 _"keeping what time has proven to be effective, and throwing out what is not"_.   

Basically, some clever guys/gals took the classic `data.frame` , shook it til the bad parts fell out, then added some clever new features. 

### _That word for needing something before hand_

```{r eval = FALSE}
#Tibbles are a part of the tidyverse package, so the easiest way to get access is to install the whole `tidyverse` package
install.packages("tidyverse")

#alternatively, install only tibble
install.packages("tibble")
```
```{r eval = FALSE}
library("tidyverse")
library("tibble")
```
```{r include = FALSE}
library("tidyverse")
library("tibble")
```

### _Tribblemaking_

There are 3 ways to make a tibble. It pretty much acts as your friendly old pal `data.frame` does.  

1) Creation. Just as `data.frame` created data frames,`tibble`, `data_frame` and `tribble` all create tibbles. 

<div class="col2">
```{r}
data.frame(a = 1:5, 
           b = letters[1:5])  
```

```{r}
data_frame(a = 1:5, 
           b = letters[1:5])
```

```{r}
tibble(a = 1:5, 
           b = letters[1:5])
```

<div>
